The Power of Unbiased Recursive Partitioning:\\ A Unifying View of CTree, MOB, and GUIDE

### 2020-02-09 pkg/partykit/inst/guideliketest/draft/draft.Rnw
further touch-ups to new paragraphs
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2978 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 523cb0dd86a3ae532f555401d0acbf36aeddddd0

\hlc{GUIDE and MOB, which are both model-based approaches, employ the model estimated 
in the considered node for an exhaustive search of the best split point over all possible 
split points in the already selected split variable.
CTree has not been developed specifically for model-based partitioning, hence, a 
different approach to selecting a split point is suggested in the original paper. 
However, as an adaption of CTree employing models for the split variable selection 
is considered in this study, it is also combined with the same exhaustive search 
strategy as GUIDE and MOB.}
%\newpage

@@

Note that there are further differences in the testing strategies between the three algorithms, e.g.,
using conditional vs.\ unconditional approximations of the null distributions. However, this difference
has relatively little influence compared to the other building blocks considered in detail here.
Moreover, both similarities and relative differences between these approaches have been previously
discussed, e.g., in \cite{Hothorn+Zeileis:2008} and \cite{Zeileis+Hothorn:2013}.
\hlc{Therefore, in this study the focus is put on the selected building blocks as they are assumed 
(and have shown) to have the highest influence on the difference in performance between the testing 
strategies.}

@@

Figure~\ref{fig:scores_stump} \hlc{also shows that for $\delta = 0$, i.e., no difference
in the parameters and consequently a random split variable selection, the probability of 
selecting $Z_1$ as split variable is (close to)~$0$ (with the employed Bonferroni-adjustment of 
$p$-values keeping this value particularly low).
Even though it has to be noted that the number of replications in this simulation study is of
course limited, the results for this setting indicate that all testing strategies conform well with 
the predefined level of significance $\alpha = 0.05$ as the probability of splitting in any 
of the available covariates is between 4 and 5~percent for $\delta = 0$ 
(the exact values can be found in} Table~\ref{tab:typeIerror} in Appendix~\ref{app:typeIerror})
\hlc{with the tests for structural changes being slightly more conservative} 
\citep[see, e.g.,][for more details on this typical behaviour]{Hothorn+Zeileis:2008, Zeileis+Hothorn:2013, Hansen:1997}.

>>>>

\hlc{Once a split variable has been selected, the selection of the corresponding
split point is less controversial. The most natural strategy is to use an exhaustive
search over all possible split points and then select the one that optimizes the
objective function (or loss function). In fact, this is the strategy adopted by both
model-based approaches, GUIDE and MOB. Only CTree uses a slightly different approach,
based on maximizing a two-sample test statistic, because the algorithm as originally
proposed did not employ an explicit objective function. However, it can be shown
that in some settings the two approaches coincide} \citep{Zeileis+Hothorn:2013}
\hlc{while the differ somewhat more generally. As we focus on model-based partitioning
here we employ the exhaustive model-based search for split point selection in
all three algorithms considered.}

@@

Note that there are further differences in the testing strategies between the three algorithms, e.g.,
using conditional vs.\ unconditional approximations of the null distributions. However, this difference
has relatively little influence compared to the other building blocks considered in detail here.
Moreover, both similarities and relative differences between these approaches have been previously
discussed, e.g., in \cite{Hothorn+Zeileis:2008} and \cite{Zeileis+Hothorn:2013}.
\hlc{Therefore, this aspect is not considered again here but focus is given to the building
blocks selected above.}

@@

Figure~\ref{fig:scores_stump} \hlc{also shows that for $\delta = 0$, i.e., no difference
in the parameters and consequently a random split variable selection, the probability of 
selecting $Z_1$ as split variable is (close to)~$0$ (with the employed Bonferroni-adjustment of 
$p$-values keeping this value particularly low).
Even with the limited number of replications in this simulation study the empirical size
of all testing strategies (i.e., for $\delta = 0$) is between 4 and 5~percent and thus
conforms well with the nominal significance level of $\alpha = 0.05$. The exact values
can be found in} Table~\ref{tab:typeIerror} in Appendix~\ref{app:typeIerror})
\hlc{which also shows that the asymptotic structural change tests are slightly more conservative}
\citep[confirming previous results in][]{Hothorn+Zeileis:2008, Zeileis+Hothorn:2013}.
